Field monitoring of Onion Maggot (\textit{Delia antiqua}) fly through improved trapping

### 2020-01-02 manuscript/journal-applied-entomology/main.tex
updated reviewer response

\begin{abstract}
Onion maggot (\textit{Delia antiqua}) is an economically important pest of Allium crops in temperate regions throughout the world.  Management of this pest is necessary to achieve economic returns and depends on insecticide regimes and cultural management.  Future management strategies for \textit{D. antiqua} may rely upon effective monitoring of \textit{D. antiqua} populations, but traditional means of monitoring can be ineffective at catching sufficient numbers of \textit{D. antiqua} flies to accurately reflect population dynamics. We evaluated the effect of shape, size, color and chemical attractants on trap catch of field populations of adult \textit{D. antiqua} flies in upstate New York. White, large diameter, spherical traps in conjunction with Delia Lure attractant performed the best in attracting and catching \textit{D. antiqua} adults. These results suggest an improved means of attracting and capturing \textit{D. antiqua} populations which could be useful in monitoring efforts and development of attract and kill strategies for pest control.

@@

\textit{D. antiqua} is a primary pest in part due to its lifecycle.  It is multivoltine with three generations per year in the Northern United States \citep{eckenrode1975population, hoepting2004insecticide}.  Larvae feed on plant roots and allium bulbs causing the damage associated with this pest.  First-generation infestations are particularly devastating as early feeding by \textit{D. antiqua} larvae can kill seedlings \citep{nault2006onion, nault2006performance}.  In addition, damaged plants become more susceptible to infestation by subsequent generations and pathogens \citep{eckenrode1986impact,nault2006performance}.  Even if plants are not completely killed by \textit{D. antiqua} larvae, feeding damage often renders the bulbs unmarketable.

@@

In situations where an insecticide application at planting is not desired for \textit{D. antiqua} control, reliance on monitoring \textit{D. antiqua} adult activity in the field can be important for deploying curative management tactics. Additionally, effective monitoring becomes even more important where \textit{D. antiqua} is a relatively recent introduction or in areas where producers are not sure they have a problem.

@@

\paragraph{Shape Trial} To evaluate effect of trap shape on trap catch and monitoring of adult \textit{D. antiqua} flies, square, cube, short cylindrical, tall cylindrical, and spherical semi-gloss white shapes each with a surface area of 180 $cm^2$  were coated with Tanglefoot sticky material (The Scotts Company LLC, Maryville, OH) and placed along the long edge of an onion (\textit{Allium cepa} L.) field near Elba, NY. Traps were spaced 15.2 meters apart and arranged in a randomized control block design with five replications.

@@

\paragraph{Shape Trial} Models relating trap shape and \textit{D. antiqua} sex to trap catch significantly explained approximately 70\% of observed variation (P \textless 0.001, $R^2_{adj}$ = 0.70, Table \ref{table:1}).  Trap shape significantly influenced (P = 0.005) catch of adult \textit{D. antiqua} (Figure \ref{fig:figure1}).  Sticky traps caught approximately 27.8 $\pm$ 2.8 more females than males.  Short cylindrical traps outperformed square panel and tall cylinder traps in terms of female trap catch.  While there were no significant differences between numbers of females captured on short cylinder, sphere, and cube traps, short cylinders tended to outperform cubes and spheres.  No significant differences were observed among treatments for male trap catch.

@@

\paragraph{Size Trial} Models using size, sex, and year to explain observed trap catch significantly explained approximately 73\% of the observed variation (P \textless 0.001, $R^2_{adj}$ = 0.73, Table \ref{table:1}).  Numbers of females captured were significantly greater than numbers of males, especially in 2005. Trap size significantly influenced trap catch (P \textless 0.001).  Increasing diameter of trap size showed a trend towards increasing trap catch (Figure \ref{fig:figure2}) with the largest trap catching significantly more females and males.  While male trap catch was less than that of females, trends in catch among the different diameters were similar for both males and females.

\paragraph{Color Trial} Models using color and sex to explain observed trap catch significantly explained approximately 79\% of the observed variation (P \textless 0.001, $R^2_{adj}$ = 0.79, Table \ref{table:1}). Color significantly influenced trap catch (P \textless 0.001) with white traps catching significantly more females than any other color (t \textless -3.1, df = 32, P \textless 0.012) and significantly more males than any other color except yellow (t \textless -3.1, df = 32, P \textless 0.012, Yellow: t = -1.6, df = 32, P = 0.3).  White traps caught significantly more females than males (t = 6.1, df = 32, P \textless 0.001).

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/publication/figure-4.pdf}
\caption{\textit{D. antiqua} fly catch per trap over the sampling period on sticky traps baited with Delia Lure attractant and unbaited.  Shaded points denote raw values while solid points and error bars denote means and bootstrapped 95\% confidence intervals respectively.  \textbf{*} indicates significantly different trap catch on baited and unbaited traps (P = 0.030). }
\label{fig:figure4}
\end{figure}

@@

Varying the shape, size and color of sticky traps revealed that changes in these factors can increase catch of \textit{D. antiqua} flies.  Short cylinder traps performed better than most other shapes, but were not significantly different from spherical traps.  Spherical traps were more economical to produce and use so they were selected for use in subsequent studies.  Size played a large role in increasing trap catch.  Larger diameter traps with more surface area captured more adult \textit{D. antiqua} flies.  Color had perhaps the largest impact on catch of adult \textit{D. antiqua} flies with white traps far outperforming yellow, green and red traps.

@@

Sticky traps have been used in monitoring and control efforts of other visually responsive flies.  Orange spheres have been effective at attracting and trapping Caribbean Fruit Fly (\textit{Anastrepha suspensa} Lowe) \citep{sivinski1990colored}.  Apple maggot (\textit{Rhagoletis pomonella} Walsh) management, in particular, has benefited from development of red spherical traps for monitoring field populations \citep{prokopy1975apple, drummond1984comparative, duan1994apple}.  This work has even expanded into a control method by trapping apple maggot flies using these spheres in conjunction with other strategies, which hasreduced infestations \citep{bostanian1999managing}.

The ability to attract larger numbers of adult \textit{D. antiqua} flies also creates an opportunity for a different, more environmentally friendly, management method. If larger numbers of flies can be attracted and then killed either by adhering to a sticky substance or coming into contact with an insecticide, populations of \textit{D. antiqua} could be reduced.  'Attract and kill' and 'Lure and kill' methodologies have been applied in pest management systems to moths, weevils, and other flies \citep{el2009potential,el2011bait,navarro2013efficacy,charmillot2000attract}.  Using large, white, spherical traps in conjunction with Delia Lure may provide the attraction needed to make it possible to control these pests using attract and kill strategies.

>>>>

\begin{abstract}
Onion maggot (\textit{Delia antiqua}) is an economically important pest of Allium crops in temperate regions throughout the world.  Management of this pest is necessary to achieve economic returns and depends on insecticide regimes and cultural management.  Current cultural management especially altering planting date, field location, and crop rotation depend on monitoring.  We evaluated the effect of shape, size, color and chemical attractants on trap catch of field populations of adult \textit{D. antiqua} flies in upstate New York. White, large diameter, spherical traps in conjunction with Delia Lure attractant performed the best in attracting and catching \textit{D. antiqua} adults. These results suggest an improved means of attracting and capturing \textit{D. antiqua} populations which could be useful in monitoring efforts and development of attract and kill strategies for pest control.

@@

\textit{D. antiqua} is a primary pest in part due to its lifecycle.  It is multivoltine with three generations per year in the Northern United States \citep{eckenrode1975population, hoepting2004insecticide}.  Larvae feed on plant roots and allium bulbs causing the damage associated with this pest.  First-generation infestations are particularly devastating as early feeding by \textit{D. antiqua} larvae can kill seedlings \citep{nault2006onion, nault2006performance}.  In addition, damaged plants become more susceptible to infestation by subsequent generations and plant pathogens \citep{eckenrode1986impact,nault2006performance}.  Even if plants are not completely killed by \textit{D. antiqua} larvae, feeding damage often renders the bulbs unmarketable.

@@

In situations where an insecticide application at planting is not desired for \textit{D. antiqua} control (such as organic management systems), reliance on monitoring \textit{D. antiqua} adult activity in the field can be important for deploying curative management tactics. Additionally, effective monitoring becomes even more important where \textit{D. antiqua} is a relatively recent introduction or in areas where producers are not sure they have a problem.

@@

\paragraph{Shape Trial} To evaluate effect of trap shape on trap catch and monitoring of adult \textit{D. antiqua} flies, square, cube, short cylindrical, tall cylindrical, and spherical semi-gloss white shapes each with a surface area of 180 $cm^2$  were coated with Tanglefoot sticky material (The Scotts Company LLC, Maryville, OH) and placed along the long edge of rectangular onion (\textit{Allium cepa} L.) fields near Elba, NY. Traps were spaced 15.2 meters apart and arranged in a randomized control block design with five replications.

@@

\paragraph{Shape Trial} Models relating trap shape and \textit{D. antiqua} sex to trap catch significantly explained approximately 70\% of observed variation (Table \ref{table:1}).  Trap shape significantly influenced (P = 0.005) catch of adult \textit{D. antiqua} (Figure \ref{fig:figure1}).  Sticky traps caught approximately 27.8 $\pm$ 2.8 more females than males.  Short cylindrical traps outperformed square panel and tall cylinder traps in terms of female trap catch.  While there were no significant differences between numbers of females captured on short cylinder, sphere, and cube traps, short cylinders tended to outperform cubes and spheres catching 10.4$\pm$ 6.2 (P = 0.3) more females.  No significant differences were observed among treatments for male trap catch.

@@

\paragraph{Size Trial} Models using size, sex, and year to explain observed trap catch significantly explained approximately 73\% of the observed variation (Table \ref{table:1}).  Numbers of females captured were significantly greater than numbers of males, especially in 2005. Trap size significantly influenced trap catch (P \textless 0.001).  Increasing diameter of trap size showed a trend towards increasing trap catch (Figure \ref{fig:figure2}) with the largest trap catching significantly more females and males.  While male trap catch was less than that of females, trends in catch among the different diameters were similar for both males and females.

\paragraph{Color Trial} Models using color and sex to explain observed trap catch significantly explained approximately 79\% of the observed variation (Table \ref{table:1}). Color significantly influenced trap catch (P \textless 0.001) with white traps catching significantly more females than any other color (t \textless -3.1, df = 32, P \textless 0.012) and significantly more males than any other color except yellow (t \textless -3.1, df = 32, P \textless 0.012, Yellow: t = -1.6, df = 32, P = 0.3).  White traps caught significantly more females than males (t = 6.1, df = 32, P \textless 0.001).

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 8cm]{figures/publication/figure-4.pdf}
\caption{\textit{D. antiqua} fly catch per trap over the sampling period on sticky traps baited with Delia Lure attractant and unbaited.  Shaded points denote total replicate values while solid points and error bars denote global means and bootstrapped 95\% confidence intervals respectively.  \textbf{*} indicates significantly different trap catch on baited and unbaited traps (P = 0.030). }
\label{fig:figure4}
\end{figure}

@@

Varying the shape, size and color of sticky traps revealed that changes in these factors can increase catch of \textit{D. antiqua} flies.  Short cylinder traps performed better than most other shapes, but were not significantly different from spherical traps.  Spherical traps were more economical to produce and use so they were selected for use in subsequent studies.  Size played a large role in increasing trap catch.  Larger diameter traps with more surface area captured more adult \textit{D. antiqua} flies.  Color had perhaps the largest impact on catch of adult \textit{D. antiqua} flies with white traps far outperforming yellow, green and red traps.  Interestingly, catch of male flies was consistently lower than catch of female flies suggesting differences in dispersion or attraction to tested designs.

@@

Sticky traps have been used in monitoring and control efforts of other visually responsive flies.  Carrot rust fly (\textit{Psila rosae}) and cabbage fly (\textit{Delia radicum}) have been monitored and managed using trapping data for over a decade \citep{stevenson1981carrot,collier1990monitoring,getzin1983carrot, broatch1997comparison, finch1990effectiveness}.   Orange spheres have been effective at attracting and trapping Caribbean Fruit Fly (\textit{Anastrepha suspensa} Lowe) \citep{sivinski1990colored}.  Apple maggot (\textit{Rhagoletis pomonella} Walsh) management, in particular, has benefited from development of red spherical traps for monitoring field populations \citep{prokopy1975apple, drummond1984comparative, duan1994apple}.  This work has even expanded into a control method by trapping apple maggot flies using these spheres in conjunction with other strategies, which has reduced infestations \citep{bostanian1999managing}.

The ability to attract larger numbers of adult \textit{D. antiqua} flies also creates an opportunity for improving current management methods in an integrated pest management context.  Current cultural management especially altering planting date, field location, and crop rotation depend on monitoring.  While trap catch alone is not a measure of monitoring efficacy, the higher numbers of \textit{D. antiqua} flies caught in these traps could potentially be used to develop better monitoring approaches.  If larger numbers of flies can be attracted to these traps, and then killed either by adhering to a sticky substance or coming into contact with an insecticide, populations of \textit{D. antiqua} may be able to be reduced.
